---
title: "How to untar and read DWD regnie data"
author: "Maximilian Nölscher"
date: "16 April 2020"
output: github_document
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(messages = FALSE)
knitr::opts_chunk$set(warnings = FALSE)
```

```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy()
```

For the following code chunks we need to load all required libraries
```{r}
library(esmisc)
library(raster)
library(tidyverse)
```


# Untar

- The DWD regnie data can be found on the [DWD ftp Server of Climate Data Center](https://opendata.dwd.de/climate_environment/CDC/grids_germany/daily/regnie/)
- Each file is named as `daily_regnie_ra<Y>m.tar`
- Each file itself contains 365 `ra<ymd>.gz` file, one per day.

For this example I just put 5 `.tar` files into the directory `../../raw_data/dwd_regnie_data/`
```{r}
directory_tar_files <- "../../raw_data/dwd_regnie_data/"
```

First we need to untar the files.
We first create a list of all files
```{r}
tar_files_list <- directory_tar_files %>% list.files()
```

Show files
```{r}
tar_files_list
```

We create a new folder to put in the untared files
```{r}
dir.create("../../processed_data/")
dir.create("../../processed_data/dwd_regnie_data/")
```


Now we untar all files. This might take a while. For 5 `.tar` files it took approx. 30 seconds.
```{r, eval=FALSE}
directory_tar_files %>% 
  paste0(tar_files_list) %>% 
  map(
    ~untar(.,
           exdir = "../../processed_data/dwd_regnie_data/")
    )
```

# Read in/ Import

Now we can start to read in the `.gz` files

First, we list all files again.
```{r}
directory_gz_files <- "../../processed_data/dwd_regnie_data/"
```

```{r}
gz_files_list <- directory_gz_files %>% list.files()
```

Show files
```{r}
gz_files_list %>% head()
```

Now comes the core chunk to read in the files. I just read in the first 5 files for demonstration
```{r}
raster_files <- directory_gz_files %>% 
  paste0(gz_files_list[1:5]) %>% 
  map(esmisc::read_regnie)
```

Plot the rasters
```{r}
raster_files %>% 
  map(plot)
```


Now you have different options
## Create Raster Stack
```{r}
raster_stack <- raster_files %>% 
  stack()
```

## ... and do calculations

e.g. create the sum
```{r}
raster_stack %>% 
  calc(sum)
```

## Reproject the CRS
```{r}
raster_files %>% 
  map(projectRaster, crs = "+init=epsg:32632")
```


## Save as GTiff

We create a new folder to put in the `.tif` files
```{r}
dir.create("../../processed_data/geotiffs/")
```

We create a vector containing the file names
```{r}
tif_file_names <- gz_files_list[1:5] %>% 
  str_remove(".gz") %>% 
  str_c("../../processed_data/geotiffs/",
        .,
        ".tif")
```

We save the rasters as `.tif` files
```{r}
raster_files %>% 
  map2(.x = .,
       .y = tif_file_names, 
       writeRaster, format="GTiff", overwrite=TRUE)
```

